---
title: "The `hubEnsembles` package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The `hubEnsembles` package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4
)
```

# Introduction

The `hubEnsembles` package includes functionality for aggregating model outputs, such as forecasts or projections, that are submitted to a hub by multiple models into combined ensemble model outputs. Currently the package includes just one function: `simple_ensemble`. We illustrate that function in this vignette.

This vignette uses the following R packages:

```{r setup}
library(dplyr)
library(plotly)
library(hubUtils)
library(hubEnsembles)
```

# Example data: a simple forecast hub

The `example-simple-forecast-hub` has been created by the Consortium of Infectious Disease Modeling Hubs as a simple example hub to demonstrate the set up and functionality for the hubverse. The hub includes both target data and example model output data.

```{r}
hub_path <- system.file("example-data/example-simple-forecast-hub",
                        package = "hubEnsembles")

model_outputs <- hubUtils::connect_hub(hub_path) %>%
  dplyr::collect()
head(model_outputs)

target_data_path <- file.path(hub_path, "target-data",
                              "covid-hospitalizations.csv")
target_data <- read.csv(target_data_path)
head(target_data)
```

# Creating ensembles with `simple_ensemble`

```{r}
mean_ens <- hubEnsembles::simple_ensemble(model_outputs)
head(mean_ens)
```

# Plots

```{r}
basic_plot_function <- function(plot_df, truth_df, plain_line = 0.5, ribbon = c(0.975, 0.025),
                                forecast_date) {

  plain_df <- dplyr::filter(plot_df, output_type_id == plain_line)

  ribbon_df <- dplyr::filter(plot_df, output_type_id %in% ribbon) %>%
    dplyr::mutate(output_type_id = ifelse(output_type_id == min(ribbon),
                                          "min", "max")) %>% 
    tidyr::pivot_wider(names_from = output_type_id, values_from = value)

  plot_model <- plot_ly(height = 600, colors = scales::hue_pal()(50)) 

  if (!is.null(truth_df)) {
    plot_model <- plot_model %>% 
      add_trace(data = truth_df, x = ~time_idx, y = ~value, type = "scatter",
                mode = "lines+markers", line = list(color = "#6e6e6e"),
                hoverinfo = "text", name = "ground truth",
                hovertext = paste("Date: ", truth_df$time_value, "<br>", 
                                  "Ground truth: ", 
                                  format(truth_df$value, big.mark = ","), 
                             sep = ""), 
                marker = list(color = "#6e6e6e", size = 7))
  }
  plot_model <- plot_model %>% 
    add_lines(data = plain_df, x = ~target_date, y = ~value, 
              color = ~model_id) %>% 
    add_ribbons(data = ribbon_df, x = ~target_date, ymin = ~min, 
                ymax = ~max, color = ~model_id, opacity = 0.25, 
                line = list(width = 0), showlegend = FALSE) %>%
    plotly::layout(shapes = list(type = "line", y0 = 0, y1 = 1, yref = "paper",
                                x0 = forecast_date, x1 = forecast_date,
                                line = list(color = "gray")))
}
```

```{r}
plot_df <- dplyr::bind_rows(model_outputs, mean_ens) %>%
  dplyr::filter(location == "US", origin_date == "2022-12-12") %>%
  dplyr::mutate(target_date = origin_date + horizon)

plot <- basic_plot_function(
    plot_df,
    truth_df = target_data %>%
        dplyr::filter(location == "US",
                      time_idx >= "2022-10-01",
                      time_idx <= "2023-03-01"),
    forecast_date = "2022-12-12")
plot
```
