---
title: "The `hubEnsembles` package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The `hubEnsembles` package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4
)
options(width = 300)
```

# Introduction

The `hubEnsembles` package provides a flexible framework for aggregating model outputs, such as forecasts or projections, that are submitted to a hub by multiple models and combined into ensemble model outputs. The package includes two main functions: `simple_ensemble` and `linear_pool`. We illustrate these functions in this vignette, and briefly compare them.

This vignette uses the following R packages:

```{r setup, message=FALSE}
library(dplyr)
library(ggplot2)
library(hubUtils)
library(hubVis)
library(hubEnsembles)
```

# Example data: a forecast hub

We will use an example hub provided by the hubverse to demonstrate the functionality of the `hubEnsembles` package. This example hub was generated with modified forecasts from the FluSight forecasting challenge, a collaborative modeling exercise run by the US Centers for Disease Control and Prevention (CDC) since 2013 that solicits seasonal influenza forecasts from outside modeling teams. The example hub includes both example model output data and target data (sometimes known as "truth" data), which are stored in the `hubExamples` package as data objects named `forecast_outputs` and `forecast_target_ts`. Note that the toy model outputs contain predictions for only a small subset rows of select dates, locations, and output type IDs, far fewer than an actual modeling hub would typically collect.

The model output data includes quantile, mean and median forecasts of future incident influenza hospitalizations and PMF forecasts of hospitalization intensity (categories determined by threshold of weekly hospital admissions per 100,000 population). Each forecast is made for five task ID variables, including the location for which the forecast was made (`location`), the date on which the forecast was made (`reference_date`), the number of steps ahead (`horizon`), the date of the forecast prediction (a combination of the date the forecast was made and the forecast horizon, `target_end_date`), and the forecast target (`target`). Below we print a subset of this example model output.

```{r example-model-outputs}
hubExamples::forecast_outputs |>
  dplyr::filter(
    output_type %in% c("quantile", "median", "pmf"),
    output_type_id %in% c(0.25, 0.75, NA, "low", "moderate", "high", "very high"),
    reference_date == "2022-12-17",
    location == "25",
    horizon == 1
  )
```

We also have corresponding target data included in the `hubExamples` package. The example target data provide observed incident influenza hospitalizations (`observation`) in a given week (`date`) and for a given location (`location`). This target data could be used as calibration data for generating forecasts or for evaluating the forecasts post hoc. The forecast-specific task ID variables `reference_date` and `horizon` are not relevant for the target data.

```{r example-target-data}
head(hubExamples::forecast_target_ts, 10)
```


# Creating ensembles with `simple_ensemble`

The `simple_ensemble()` function directly computes an ensemble from component model outputs by combining them via some function within each unique combination of task ID variables, output types, and output type IDs. This function can be used to summarize predictions of output types mean, median, quantile, CDF, and PMF. The mechanics of the ensemble calculations are the same for each of the output types, though the resulting statistical ensembling method differs for different output types.

By default, `simple_ensemble()` uses the mean for the aggregation function and equal weights for all models, though the user can create different types of weighted ensembles by specifying an aggregation function and weights. 

Using the default options for `simple_ensemble()`, we can generate an equally weighted mean ensemble for each unique combination of values for the task ID variables, the `output_type` and the `output_type_id`. This means different ensemble methods will be used for different output types: for the `quantile` output type in our example data, the resulting ensemble is a quantile average, while for the mean, CDF, PMF output type, the ensemble is a linear pool. We must filter the sample output type because it is not yet supported.

```{r}
mean_ens <- hubExamples::forecast_outputs |>
  dplyr::filter(output_type != "sample") |>
  hubEnsembles::simple_ensemble(
    model_id = "simple-ensemble-mean"
  )
```

The resulting model output has the same structure as the original model output data, with columns for model ID, task ID variables, output type, output type ID, and value. We also use `model_id = "simple-ensemble-mean"` to change the name of this ensemble in the resulting model output; if not specified, the default will be "hub-ensemble". A subset of the predictions is printed below.

```{r mean-ensemble}
mean_ens |>
  dplyr::filter(
    output_type %in% c("quantile", "median", "pmf"),
    output_type_id %in% c(
      0.025, 0.25, 0.75, 0.975, NA,
      "low", "moderate", "high", "very high"
    ),
    reference_date == "2022-12-17",
    location == "25",
    horizon == 1
  )
```

## Changing the aggregation function

We can change the function that is used to aggregate model outputs. For example, we may want to calculate a median of the component models' submitted values for each quantile. We do so by specifying `agg_fun = median`.

```{r}
median_ens <- hubExamples::forecast_outputs |>
  dplyr::filter(output_type != "sample") |>
  hubEnsembles::simple_ensemble(
    agg_fun = median,
    model_id = "simple-ensemble-median"
  )
```

Custom functions can also be passed into the `agg_fun` argument. We illustrate this by defining a custom function to compute the ensemble prediction as a geometric mean of the component model predictions. Any custom function to be used must have an argument `x` for the vector of numeric values to summarize, and if relevant, an argument `w` of numeric weights.

```{r}
geometric_mean <- function(x) {
  n <- length(x)
  return(prod(x)^(1 / n))
}
geometric_mean_ens <- hubExamples::forecast_outputs |>
  dplyr::filter(output_type != "sample") |>
  hubEnsembles::simple_ensemble(
    agg_fun = geometric_mean,
    model_id = "simple-ensemble-geometric"
  )
```

As expected, the mean, median, and geometric mean each give us slightly different resulting ensembles. The median point estimates, 50% prediction intervals, and 90% prediction intervals in the figure below demonstrate this. Note that the geometric mean ensemble and simple mean ensemble generate similar estimates in this case of predicting weekly incident influenza hospitalizations in Massachusetts.

```{r plot-ensembles, echo=FALSE, warning=FALSE, fig.height=4, fig.width=8}
model_out_plot <- dplyr::bind_rows(
  mean_ens, median_ens,
  geometric_mean_ens
) |>
  dplyr::filter(
    location == "25",
    output_type %in% c("mean", "quantile"),
    reference_date == "2022-12-17"
  ) |>
  dplyr::mutate(target_date = reference_date + horizon)

target_data_plot <- hubExamples::forecast_target_ts |>
  dplyr::filter(
    location == "25",
    date >= "2022-11-01", date <= "2023-03-01"
  )

hubVis::plot_step_ahead_model_output(
  model_output_data = model_out_plot,
  target_data = target_data_plot,
  use_median_as_point = TRUE,
  interactive = FALSE,
  intervals = c(0.5, 0.90),
  show_legend = TRUE,
  x_col_name = "target_end_date",
  x_target_col_name = "date"
) +
  theme_bw() +
  labs(y = "incident hospitalizations")
```

## Weighting model contributions

We can weight the contributions of each model in the ensemble using the `weights` argument of `simple_ensemble()`. This argument takes a `data.frame` that should include a `model_id` column containing each unique model ID and a `weight` column. In the following example, we include the baseline model in the ensemble, but give it less weight than the other forecasts.

```{r}
model_weights <- data.frame(
  model_id = c("MOBS-GLEAM_FLUH", "PSI-DICE", "simple_hub-baseline"),
  weight = c(0.4, 0.4, 0.2)
)
weighted_mean_ens <- hubExamples::forecast_outputs |>
  dplyr::filter(output_type != "sample") |>
  hubEnsembles::simple_ensemble(
    weights = model_weights,
    model_id = "simple-ensemble-weighted-mean"
  )
head(weighted_mean_ens, 10)
```

# Creating ensembles with `linear_pool`

The `linear_pool()` function implements the linear opinion pool (LOP, also known as a distributional mixture) method when ensembling predictions. This function can be used to combine predictions with output types mean, quantile, CDF, and PMF. Unlike `simple_ensemble()`, this function handles its computation differently based on the output type. For the CDF, PMF, and mean output types, the linear pool method is equivalent to calling `simple_ensemble()` with a mean aggregation function, since `simple_ensemble()` produces a linear pool prediction (an average of individual model cumulative or bin probabilities).

For the quantile output type, the `linear_pool()` function first must approximate a full probability distribution using the value-quantile level pairs from each component model. As a default, this is done with functions in the `distfromq` package, which defaults to fitting a monotonic cubic spline for the interior and a Gaussian normal distribution for the tails. Quasi-random samples are drawn from each distributional estimate, which are then collected and used to extract the desired quantiles from the final ensemble distribution.

Using the default options for `linear_pool()`, we can generate an equally-weighted linear pool for each of the output types in our example hub (except for the median and sample output types, which must be excluded). The resulting distribution for the linear pool of quantiles is estimated using a default of `n_samples = 1e4` quasi-random samples drawn from the distribution of each component model.

```{r}
linear_pool_norm <- hubExamples::forecast_outputs |>
  dplyr::filter(!output_type %in% c("median", "sample")) |>
  hubEnsembles::linear_pool(model_id = "linear-pool-normal")
head(linear_pool_norm, 10)
```

In the figure below, we compare ensemble results generated by `simple_ensemble()` and `linear_pool()` for model outputs of output types PMF and quantile. Panel A shows PMF type predictions of Massachusetts incident influenza hospitalization intensity while Panel B shows quantile type predictions of Massachusetts weekly incident influenza hospitalizations. As expected, the results from the two functions are equivalent for the PMF output type: for this output type, the `simple_ensemble()` method averages the predicted probability of each category across the component models, which is the definition of the linear pool ensemble method. This is not the case for the quantile output type, because the `simple_ensemble()` is computing a quantile average.

```{r plot-ex-quantile-and-linear-pool, echo=FALSE, fig.height=4, fig.width=10}
# Generate plot of ensemble probability in each bin
p1 <- dplyr::bind_rows(mean_ens, linear_pool_norm) |>
  dplyr::filter(
    output_type == "pmf", reference_date == "2022-12-17",
    location == "25"
  ) |>
  # Reorder bin names to be in correct order
  dplyr::mutate(
    output_type_id =
      factor(
        output_type_id,
        levels = c("low", "moderate", "high", "very high")
      )
  ) |>
  ggplot(aes(x = output_type_id, y = value, fill = model_id)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(vars(horizon), labeller = label_both) +
  labs(x = "incident hospitalization intensity", y = "probability") +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  theme_bw() +
  theme(
    legend.position = "bottom", legend.title = element_blank(),
    strip.background = element_blank(), strip.placement = "outside",
    panel.grid.major.x = element_blank()
  )

## Create new data frame of ensemble output to be plotted
# Use median results as point estimate in plot
# Pull quantile == 0.5 results for linear pool and add them to the end of
# the data.frame as "median" output type (because linear_pool() does not
# automatically provide median estimates)
model_output_plot <- linear_pool_norm |>
  dplyr::filter(output_type_id == 0.5, target == "wk inc flu hosp") |>
  dplyr::mutate(output_type = "median", output_type_id = NA)
model_output_plot <- dplyr::bind_rows(linear_pool_norm, model_output_plot)
# Filter to desired forecasts
model_output_plot <- dplyr::bind_rows(mean_ens, model_output_plot) |>
  dplyr::filter(
    location == "25",
    output_type %in% c("median", "mean", "quantile"),
    reference_date == "2022-12-17"
  ) |>
  dplyr::mutate(
    target_date = reference_date + horizon,
    output_type_id = as.numeric(output_type_id)
  )
target_data_plot <- hubExamples::forecast_target_ts |>
  dplyr::filter(
    location == "25",
    date >= "2022-11-01", date <= "2023-03-01"
  )

# Generate plot of incidence forecasts 0-3 week horizons
p2 <-
  hubVis::plot_step_ahead_model_output(
    model_output_data = model_output_plot,
    target_data = target_data_plot,
    use_median_as_point = TRUE,
    interactive = FALSE,
    intervals = c(0.5, 0.9),
    pal_color = "Set1",
    show_legend = TRUE,
    x_col_name = "target_end_date",
    x_target_col_name = "date"
  ) +
  theme_bw() +
  labs(y = "incident hospitalizations")

# Pull legend
l <- cowplot::get_plot_component(p1, "guide-box", return_all = TRUE)[[3]]

# Assemble plots
cowplot::plot_grid(
  cowplot::plot_grid(
    p1 + labs(subtitle = "example PMF output type") +
      theme(legend.position = "none"),
    p2 + labs(subtitle = "example quantile output type") +
      theme(legend.position = "none"),
    labels = LETTERS[1:2]
  ), l,
  ncol = 1,
  rel_heights = c(0.95, 0.05)
)
```

Like with `simple_ensemble()`, we can change the default aggregation settings. For example, weights that determine a model's contribution to the resulting ensemble may be provided.

```{r}
model_weights <- data.frame(
  model_id = c("MOBS-GLEAM_FLUH", "PSI-DICE", "simple_hub-baseline"),
  weight = c(0.4, 0.4, 0.2)
)
weighted_linear_pool_norm <- hubExamples::forecast_outputs |>
  dplyr::filter(!output_type %in% c("median", "sample")) |>
  hubEnsembles::linear_pool(
    weights = model_weights,
    model_id = "linear-pool-weighted"
  )
head(weighted_linear_pool_norm, 10)
```

We can also change the distribution that `distfromq` uses to estimate the component models' distributional tails to either log normal or cauchy using the `tail_dist` argument. (See documentation in the `distfromq` package for more details and function options.) Note, though, that the choice of tail distribution usually does not have a large impact on the resulting ensemble distribution, except in its outer edges.

```{r}
linear_pool_lnorm <- hubExamples::forecast_outputs |>
  dplyr::filter(!output_type %in% c("median", "sample")) |>
  hubEnsembles::linear_pool(
    model_id = "linear-pool-lognormal",
    tail_dist = "lnorm"
  )
linear_pool_cauchy <- hubExamples::forecast_outputs |>
  dplyr::filter(!output_type %in% c("median", "sample")) |>
  hubEnsembles::linear_pool(
    model_id = "linear-pool-cauchy",
    tail_dist = "cauchy"
  )
```

