---
title: "The `hubEnsembles` package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The `hubEnsembles` package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4
)
```

# Introduction - Emily

Predictions of future outcomes are essential to planning and decision making. Despite their utility, generating reliable predictions of the future is challenging. One method for overcoming this challenge is combining predictions across multiple, independent models. These combination methods (also called aggregation or ensembling) have been repeatedly shown to produce predictions that are more accurate [@clemen1989; @timmermann2006b] and more consistent [@hibon2005] than individual models. Because of the clear performance benefits, multi-model ensembles are commonplace across fields, including weather [@alley2019], climate [@tebaldi2007], and economics [@aastveit2018]. More recently, multi-model ensembles have been used to improve predictions of infectious disease outbreaks, including for respiratory viruses (influenza [@mcgowan2019], SARS-CoV-2 [@cramer2022]), vector borne diseases (Dengue [@johansson2019]), and hemorrhagic fevers (Ebola [@viboud2018]). <!--#EH: I DON'T LOVE THE EMPHASIS ON DISEASE APPLICATIONS… THOUGHTS ON WHETHER TO LEAVE THIS SENTENCE, SIMPLIFY IT, OR CUT IT?-->
<!--# LS: We could include infectious diseases in the list of different fields that use ensembles, then move this sentence further down to around the mention of the hubverse since its goal and our main expertise lies in infectious disease forecasting.-->

Across this vast literature, there are many proposed methods for generating ensembles. Generally, these methods differ in at least one of two ways: (1) the function used to combine or "average" predictions, and (2) how predictions are weighted when performing the combination. No one method is universally "the best"; a simple average of predictions works surprisingly well across a range of settings [@winkler2015], but more complex approaches have also been shown to have benefits [REF]. Here, we present the `hubEnsembles` package, which provides a flexible framework for generating ensemble predictions from multiple models. Complimenting an existing R package for generating ensembles of point estimates [@weiss2019], `hubEnsembles` supports multiple types of predictions, from point estimates to probabilistic predictions <!--#EH: DOES ANYONE KNOW OF OTHER PACKAGES?-->.

The `hubEnsembles` package is part of a larger collection of open-source software and data tools that enables collaborative forecasting exercises called the "hubverse". The broader "hubverse" initiative is motivated by the demonstrated benefits of collaborative hubs [@reich2022], including performance improvements of multi-model ensembles. In this paper, we focus specifically on the functionality encompassed in `hubEnsembles`. We provide an overview of the methods implemented (Section 2), give simple examples to demonstrate the functionality (Section 3) and a more complex case study (Section 4) that motivates a discussion and comparison of the various methods (Section 5).

<!--#EH GENERAL COMMENT: IF YOU HAVE BETTER REFERENCES ANYWHERE, LET ME KNOW.-->.

# Ensembling methods - Li

## Introduction to hubverse terminology and conventions
- Define: Round, Target, Task, Model Output, Nowcast, Forecast
  - Modeling tasks: task ID variables, output types, (target metadata)
  - Model output representation: output_type, output_type_id, value

## General definitions and ensemble properties
- Introduce mathematical notation
- Note properties shared by both ensembles (e.g. combining component forecasts over unique combination of task ID variables)
- Option for weighting

## Simple ensemble

## Linear pool
### CDF, PMF, and Mean output types
### Quantile output type

# Demonstration of functionality - Emily

The `hubEnsembles` package includes functionality for aggregating model outputs, such as forecasts or projections, that are submitted to a hub by multiple models and combined into ensemble model outputs. The package includes two main functions: `simple_ensemble` and `linear_pool`. We illustrate these functions in this vignette, and briefly compare them.

This vignette uses the following R packages:

```{r setup}
library(dplyr)
library(plotly)
library(hubUtils)
library(hubEnsembles)
```

# 

## Example data: a simple forecast hub

The `example-simple-forecast-hub` has been created by the Consortium of Infectious Disease Modeling Hubs as a simple example hub to demonstrate the set up and functionality for the hubverse. The hub includes both target data and example model output data.

```{r}
hub_path <- system.file("example-data/example-simple-forecast-hub",
                        package = "hubEnsembles")

model_outputs <- hubUtils::connect_hub(hub_path) %>%
  dplyr::collect()
head(model_outputs)

target_data_path <- file.path(hub_path, "target-data",
                              "covid-hospitalizations.csv")
target_data <- read.csv(target_data_path)
head(target_data)
```

## Creating ensembles with `simple_ensemble`

The `simple_ensemble` function is used to summarize across component model outputs; this function can be applied to predictions with an `output_type` of `mean`, `median`, `quantile`, `cdf`, or `pmf`.

The `simple_ensemble` function defaults to calculating an equally weighted mean across all component model outputs for each unique `output_type_id`. For our example data, which contains two output types (`median` and `quantile`), this means the resulting ensemble will be the mean of component model submitted values for each quantile.

```{r}
mean_ens <- hubEnsembles::simple_ensemble(model_outputs)
head(mean_ens)
```

### Changing the aggregation function

We can change the function used to aggregate across model outputs. For example, we may want to calculate a median of component model submitted values for each quantile. We will also use the `model_id` argument to distinguish this ensemble.

```{r}
median_ens <- hubEnsembles::simple_ensemble(model_outputs, 
                                            agg_fun = median, 
                                            model_id = "hub-ensemble-median")
head(median_ens)
```

Custom functions can also be passed into the `agg_fun` argument. For example, a geometric mean may be a more appropriate way to combine component model outputs. Any custom function to be used requires an argument `x` for the vector of numeric values to summarize, and if relevant, an argument `w` of numeric weights.

```{r}
geometric_mean <- function(x){
    n <- length(x)
    return(prod(x)^(1/n))
}

geometric_mean_ens <-  hubEnsembles::simple_ensemble(model_outputs, 
                                            agg_fun = geometric_mean, 
                                            model_id = "hub-ensemble-geometric")
head(geometric_mean_ens)
```

### Weighting model contributions

In addition, we can weight the contributions of each model by providing a table of weights, which are provided in a `data.frame` with a `model_id` column and a `weight` column.

```{r}
model_weights <- data.frame(model_id = c("UMass-ar", "UMass-gbq", "simple_hub-baseline"), 
                            weight = c(0.4, 0.4, 0.2))

weighted_mean_ens <- hubEnsembles::simple_ensemble(model_outputs, 
                                                   weights = model_weights, 
                                                   model_id = "hub-ensemble-weighted-mean")
head(weighted_mean_ens)
```

## Creating ensembles with `linear_pool`

An alternative approach to generate an ensemble is a linear pool, or a distributional mixture; this function can be applied to predictions with an `output_type` of `mean`, `quantile`, `cdf`, or `pmf`. Our example hub includes `median` output type, so we exclude it from the calculation.

For `mean`, `cdf` and `pmf` output types, the linear pool is equivalent to using a mean `simple_ensemble`. For `quantile` model outputs, the `linear_pool` function needs to approximate a full probability distribution using the value-quantile pairs from each component model. As a default, this is done with functions in the `distfromq` package, which defaults to fitting a monotonic cubic spline.

```{r}
linear_pool_ens <- hubEnsembles::linear_pool(model_outputs %>%
                                               filter(output_type != "median"))
head(linear_pool_ens)
```

## Plots

```{r}
basic_plot_function <- function(plot_df, truth_df, plain_line = 0.5, ribbon = c(0.975, 0.025),
                                forecast_date) {

  plain_df <- dplyr::filter(plot_df, output_type_id == plain_line)

  ribbon_df <- dplyr::filter(plot_df, output_type_id %in% ribbon) %>%
    dplyr::mutate(output_type_id = ifelse(output_type_id == min(ribbon),
                                          "min", "max")) %>% 
    tidyr::pivot_wider(names_from = output_type_id, values_from = value)

  plot_model <- plot_ly(height = 600, colors = scales::hue_pal()(50)) 

  if (!is.null(truth_df)) {
    plot_model <- plot_model %>% 
      add_trace(data = truth_df, x = ~time_idx, y = ~value, type = "scatter",
                mode = "lines+markers", line = list(color = "#6e6e6e"),
                hoverinfo = "text", name = "ground truth",
                hovertext = paste("Date: ", truth_df$time_value, "<br>", 
                                  "Ground truth: ", 
                                  format(truth_df$value, big.mark = ","), 
                             sep = ""), 
                marker = list(color = "#6e6e6e", size = 7))
  }
  plot_model <- plot_model %>% 
    add_lines(data = plain_df, x = ~target_date, y = ~value, 
              color = ~model_id) %>% 
    add_ribbons(data = ribbon_df, x = ~target_date, ymin = ~min, 
                ymax = ~max, color = ~model_id, opacity = 0.25, 
                line = list(width = 0), showlegend = FALSE) %>%
    plotly::layout(shapes = list(type = "line", y0 = 0, y1 = 1, yref = "paper",
                                x0 = forecast_date, x1 = forecast_date,
                                line = list(color = "gray")))
}
```

```{r}
plot_df <- dplyr::bind_rows(model_outputs, mean_ens) %>%
  dplyr::filter(location == "US", origin_date == "2022-12-12") %>%
  dplyr::mutate(target_date = origin_date + horizon)

plot <- basic_plot_function(
    plot_df,
    truth_df = target_data %>%
        dplyr::filter(location == "US",
                      time_idx >= "2022-10-01",
                      time_idx <= "2023-03-01"),
    forecast_date = "2022-12-12")
plot
```

# Case study: Weekly incident flu hospitalizations - Li

# Conclusion/Discussion - Undecided
